---
title: "The Sopranos Episodes You Should Just Skip"
output: github_document
---



```{r libraries, include=FALSE}
library(tidyverse)
library(ggrepel)
library(extrafont)
library(car)
```


```{r functions, include=FALSE}
nyt_theme <- function() {
  theme_bw() +   
  theme(
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank(),

        panel.grid.major.y = element_line(color = "#E2E2E2", size= .25),
        
        plot.title=element_text(vjust=1.25,hjust=.5),
        axis.title.x=element_text(vjust=0),
        axis.title.y=element_text(vjust=1.25)
       ) 
}

xfree_theme <- function() {
  theme(
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major.x = element_blank()
        )  
}

#stretches (or squashes) sequence in vector, 
#so that last value is first value plus 12
stretchit <- function(x) {
  #makes a sequence from 1:13 the length of given vector
  temp<-seq(from=1,
            to=13,
            length.out=length(x))
  
  #shift it to starting value of given vector 
  distance<-x[1]-temp[1]
  temp<-temp+distance
  
  return(temp)
}
```


```{r font imports, include=FALSE}
#https://stackoverflow.com/questions/34522732/changing-fonts-in-ggplot2
#uncomment next line. just run one time on your system to import fonts from specified folder to R folder
#if you want to import from your windows folder just do font_import()
font_import(path="C:/Users/Fred/Dropbox/fonts")
loadfonts(device = "win")
```



```{r read, include=FALSE}
thesopranos <- read_csv("thesopranos.csv")
thesopranos$sequential <- 1:length(thesopranos$Season)

#seasons 6.1 and 6.2 have different #s of episodes than other seasons. this spaces them out to they are the same width as the other seasons when plotting
thesopranos$sequential.spaced <- thesopranos$sequential
thesopranos[(thesopranos$Season==6.2)&(thesopranos$Episode==13),]$sequential.spaced<-79
thesopranos[(thesopranos$Season==6.1),]$sequential.spaced<-stretchit(thesopranos[(thesopranos$Season==6.1),]$sequential.spaced)
thesopranos[(thesopranos$Season==6.2),]$sequential.spaced<-stretchit(thesopranos[(thesopranos$Season==6.2),]$sequential.spaced)

thesopranos$Season <- as.factor(thesopranos$Season)
thesopranos$z.score <- (thesopranos$Rating - mean(thesopranos$Rating))/sd(thesopranos$Rating)

#position for "Season 1" text
df.summary <- thesopranos %>% group_by(Season) %>% summarise (mean.ep=mean(Episode))
df.summary$position <- seq(from=7,
                           to=85,                           
                           length.out=length(df.summary[[1]]))

df.summary$seasonlabel <- paste("SEASON ", df.summary$Season)

```



I am rewatching The Sopranos, just in time for the 20 year anniversary. It's always been one of my top TV series ever, but a lot of time has passed. I was worried that it wouldn't hold up. But it's so good, right from the very first episode. It holds up. It might even be better, funnier, and smarter than I realized when I watched it for the first time.

Then I got to the episode "A Hit is a Hit". It's just bad. Meandering plot structure, poor editing, and bad dialogue. I didn't even want to finish it.


It recieved an 8.3 on IMDb viewer ratings. I was shocked. **8.3 is not a score for a bad hour of television**. 8.3 should be, objectively, pretty good. 8.3 is something you would recommend and is worth your time. "A Hit is a Hit" is neither of those things. And, objectively, the ratings of the other episodes I watched so far weren't all that different, from 8.6 to 8.9:
<br><br>

```{r show some ratings, echo=FALSE}
thesopranos %>%
  select(Combined, EpisodeName, Rating) %>%
  filter(Combined<111) %>%
  rename(Episode = Combined,
         `Episode Name` = EpisodeName) 
```

<br>
A difference of .6 shouldn't be the gap between a bad episode and a really good episode. This is where some statistical analysis can be useful. Here is the the distribution of IMDB viewer rating scores of all the episodes:
<br>

```{r histogram, echo=FALSE, fig.width=9, fig.height=2, fig.align="center"}
ggplot(data=thesopranos, aes(x=Rating))+
  geom_histogram(binwidth=.05,
                 fill="#F8766D")+
  labs(title="IMBD Viewer Ratings of The Sopranos Episodes",
       y="# of Episodes") +
  nyt_theme() +
  scale_x_continuous(breaks=seq(8, 10, by=.1))+
  scale_y_continuous(breaks=seq(0, 18, by=4))+
  geom_hline(yintercept=0, size=0.4, color="black")+
  theme(panel.grid.major.x = element_blank())
```


This distribution is roughly [normal](https://en.wikipedia.org/wiki/Normal_distribution) enough for my purpose, which is to make sure I never waste my time watching an episode as bad as "A Hit is a Hit" again. That particular episode was 1.25 standard deviations (SD) lower than the mean rating of all episodes, so I decided to use 1 SD lower than the mean as my cutoff for "bad" episodes. In a perfectly normal dstribution, this means I wouldn't watch 17% of all the episodes. I should remind you that The Sopranos is 86 hours of television. I can live with watching 17% less of it.

##IMDb ratings visualizations

How do I visualize IMDb ratings? I found some nice plots from [u/Pyrolamas](https://www.reddit.com/r/dataisbeautiful/comments/6rzoqw/the_hillenburg_effect_spongebob_squarepants_imdb/) and [u/shivasprogeny](https://www.reddit.com/r/dataisbeautiful/comments/1g7jw2/seinfeld_imdb_episode_ratings_oc/) on reddit that seemed like a good models for me to emulate.

Here's the result. 


```{r reddit inspired, echo=FALSE, fig.width=9, fig.height=3, fig.align="center"}
ggplot(data=thesopranos, aes(x=sequential.spaced, y = Rating, color=Season, label=EpisodeName))+
  
  #data points
  geom_point(
              aes(color=Season)
             )+
  
  #curve and confidence interval
  stat_smooth(aes(group=Season),
               method="loess",
              size=1,
              alpha=.2)+
  
  #"Season X" labels
  geom_text(data=df.summary, aes(label=seasonlabel, 
                                 x=position,
                                 y=9.5),
            family="Roboto Condensed", size=4)+
labs(title="IMDB VIEWER RATINGS FOR THE SOPRANOS")+
  
  xlab("Episode")+
  scale_y_continuous(
                     breaks=c(0.0:10.0),
                     limits=c(7.25,10.0))+
  scale_fill_brewer(palette="Dark2")+
  nyt_theme()+
  xfree_theme()+
  theme(
        legend.position="none",
        plot.title=element_text(size=15,
                                hjust=0.5, 
                                family="League Spartan", 
                                color="#9A9A9A" )
  )

```

Of course, I wanted to apply my criteria for "bad episodes," then highlight and label them. Here is the final product:
<br><br><br>

```{r low points highlighted, echo=FALSE, fig.width=9, fig.height=4}

ggplot(data=thesopranos, aes(x=sequential.spaced, y = Rating, color=Season, label=EpisodeName))+
  
  #the data points
  geom_point(data=filter(thesopranos,
                         z.score > -1),
             color="gray90",
             size=1
             ) +
  
  #the curves 
  stat_smooth(
  aes(group=Season),

            method="loess",
            size=1,
            alpha=.1)+

  #horizontal line for 1sd below mean
  geom_hline(yintercept=mean(thesopranos$Rating)-sd(thesopranos$Rating),
            color="gray 70",
            linetype="dashed") +

  #labels for the bad episodes
  geom_text_repel(
    data=filter(thesopranos,
                       z.score < -1),
    color="black",
    nudge_y      = -50,
    direction    = "x",
    force=.1,
    angle        = 90,
    hjust =0.5,
    segment.size = 0.2,
    show.legend=FALSE,
    family="Roboto Condensed",
    size = 3    
    )+
  
  #highlighted points of the bad episodes
  geom_point(data=filter(thesopranos,
                       z.score < -1),
            aes(color=Season),
               color="red",
            size=2,
            alpha=.4
           ) +
  
  #"Season X" labels
  geom_text(data=df.summary, aes(label=seasonlabel, 
                                 x=position,
                                 y=9.5),
            family="Roboto Condensed", size=4.5)+
  
  #text for 1sd below mean
  geom_text(label="1 STANDARD DEVIATION BELOW THE MEAN",
            x=26, y=mean(thesopranos$Rating)-sd(thesopranos$Rating),
            family="Roboto Condensed", size=3, color="#9A9A9A",
            hjust=.5,vjust=1.9)+

  labs(title="THE SOPRANOS EPISODES YOU SHOULD JUST SKIP",
       subtitle="Episodes rated lower than 1 standard deviation below the mean on IMDB")+
  scale_y_continuous(
                     breaks=c(0.0:10.0),
                     limits=c(6,10.0))+
  scale_fill_brewer(palette="Dark2")+
  nyt_theme()+
  xfree_theme()+
  theme(panel.grid.major.y = element_blank(),
        legend.position="none",
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        plot.title=element_text(size=15,hjust=0.5, family="League Spartan", color="#9A9A9A" ),
        plot.subtitle=element_text(size=8,hjust=0.5, color="gray60")
  )



```


```{r SAVE, include=FALSE}
savescalemultiplier=1
ggsave(filename = "graphic.png", width=10*savescalemultiplier, height=4*savescalemultiplier, dpi=300, type = "cairo")
```

```{r normality tests, include=FALSE}
shapiro.test(thesopranos$Rating)
qqPlot(thesopranos$Rating)
```



```{r BOXPLOT, include=FALSE}
ggplot(data=thesopranos, aes(x=Season, y = Rating, fill=Season,color=Season))+
  geom_jitter(aes(color=Season),
              shape = 21,
              color="gray85",
              width=.1,
              height= .05,
              size=4,
              stroke=1,
              fill="white",
              alpha=1)+
  scale_shape_manual(values=c(23))+
  geom_boxplot(
               alpha=.1,
               outlier.shape = NA)+ #no outliers
  stat_summary(fun.y="mean",fun.ymin = "min", fun.ymax = "max",
               # color="gray50",
               size=4,
               geom="point",
               shape = 18)+

  theme_bw() +   
  theme(
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "#E2E2E2",size=.5)
        # panel.background = element_rect(fill = "white", colour = "grey50")
        )  
```


```{r ggbetween stats, include=FALSE}
# ggbetweenstats(data=thesopranos, x=Season, y = Rating)+
#   theme_bw() +   
#   theme(
#         panel.grid.minor = element_blank(),
#         panel.border = element_blank(),
#         axis.ticks = element_blank(),
#         axis.text.x = element_blank(),
#         axis.title.x = element_blank(),
#         panel.grid.major.x = element_blank(),
#         panel.grid.major.y = element_line(color = "#E2E2E2",size=.5)
#         # panel.background = element_rect(fill = "white", colour = "grey50")
#         )    
```

